---
title: 'BMI551 Kaggle: EDA'
author: "Nathaniel Evans"
date: "March 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # all the good stuff
library(readxl) # for reading in xlsx files
library(janitor) # for clean_names
library(knitr) # for kable
library(moderndive) # for getting tables
library(corrr) # for correlation matrix
library(skimr) # for skim
library(GGally) # for ggpairs
library(broom) # for pulling out model results
library(dplyr)
library(ggpubr)
library(randomForest)

select <- dplyr::select


library(tidyr)
library(dplyr) 
library(ggplot2) 
library(moments)
library(Rtsne)


```

# Data in 

File descriptions

`expression.txt` - tab-delimited text file with expression values for 18,632 genes for each of the 39 cell lines.
`subtypes.txt` - tab-delimited text file of subtypes (basal, luminal, claudin-low and normal-like) for each of 39 cell lines
`training_set_answers.txt` - tab-delimited text file of the correct classification of 0 (non-responsive) or 1 (responsive) for each combination of 25 cell lines and 12 drugs.
`scoring_and_test_set_id_mappings.csv` - comma-delimited text file of the id used by Kaggle for each of the cell line/drug combinations in the scoring set and test set. The first 108 values are the scoring set (9 cell lines and 12 drugs) and the last 60 are the final test set (5 cell lines 12 drugs). Scores on the final test set will not be shown until the competition is over.
`rand_sub_cont.csv` - a sample submission file in the correct format with random predictions between 0 and 1. The calculation of the AUROC value summarizes the performance of these guesses at all thresholds between 0 and 1.

```{r, message=F, warning=F}
exp.dat <- read.csv2('./data/expression.txt', header=T, sep='\t', numerals='warn.loss', stringsAsFactors = F)

rnd.sub.cnt <- read.csv2('./data/rand_sub_cont.csv', sep=',')

id.map <- read.csv2('./data/scoring_and_test_set_id_mappings.csv', sep=',')

sub.types <- read.csv2('./data/subtypes.txt', sep='\t')

train.dat <- read.csv2('./data/training_set_answers.txt', sep='\t')

gene.names <- rownames(exp.dat)
line.names <- colnames(exp.dat)
rownames(exp.dat) <- NULL
exp.dat <- mutate_all((exp.dat), function(x) as.numeric(x))
colnames(exp.dat) <- NULL
exp.dat <- t(exp.dat)
colnames(exp.dat) <- gene.names
exp.dat <- cbind(data.frame('cell.line'=line.names), exp.dat) 

#if (exp.dat$cell.line == train.dat$cell.line) { 
#  print ('there is a mismatch')  
#}

# Example Submission File: randomly chosen 
rnd.sub.cnt <- rnd.sub.cnt %>% mutate('sub.cont'=value) %>% select(-value)

# Submission file id order 
id.map <- id.map %>% mutate(cell.line = cellline) %>% select(-cellline)

# cell line subtype s
sub.types <- sub.types %>% mutate(cell.line = cellline) %>% select(-cellline)

train.dat <- cbind(data.frame('cell.line'=rownames(train.dat)), train.dat)
rownames(train.dat) <- NULL

train.dat <- train.dat %>% merge(sub.types, by='cell.line') #%>% merge(exp.dat, by='cell.line')

head(train.dat)

save(train.dat, file='./data/train.rdat')
save(exp.dat, file='./data/exp.rdat')
save(id.map, file='./data/idmap.rdat')
```

# EDA 

## skim 

```{r}
skim(train.dat)
skim()
```

## Variable Variance 

```{r}

skew <- numeric(length=length(colnames(exp.dat)))
kurt <- numeric(length=length(colnames(exp.dat)))
variance <- numeric(length=length(colnames(exp.dat)))
for (i in 2:length(colnames(exp.dat))) {
  variance[i] <- var(exp.dat[,i])
  skew[i] <- skewness(exp.dat[,i])
  kurt[i] <- kurtosis(exp.dat[,i])
}

s1 <- data.frame(variance, skew, kurt)

s1 %>% summary()
s1 <- s1 %>% filter (variance <3)
s1 %>% ggplot(aes(x=variance)) + geom_density()
s1 %>% ggplot(aes(x=skew)) + geom_density()
s1 %>% ggplot(aes(x=kurt)) + geom_density()

```

## tsne clustering 

To look at how the data clusters, I first filtered by variance then used tsne to embed in 2d space. The results show significant seperation by subtype. 

```{r}
# filter by variance 
exp.hivar <- exp.dat[(s1$variance > 3)]

#tsne
tsne <- Rtsne(exp.hivar, dims = 2, perplexity=12, verbose=TRUE, max_iter = 1000, theta=0.25)
x1 <- tsne$Y[,1]
x2 <- tsne$Y[,2]
t2 <- exp.dat %>% select(cell.line) %>% cbind(x1,x2) %>% merge(sub.types, by='cell.line')

#plot 
t2 %>% ggplot(aes(x=x1, y=x2, color=subtype)) + geom_point()
```


# heat map

Compare drug sensitivities 

```{r}

heatmap(as.matrix(exp.hivar[,-1]))


```






# feature selection

### Select top 1000 highest variance genes 

### Apply PLS on top 1000 

### Build model for each drug 


# ALternative idea 

### Treat each label type seperately
- Apply PLS to data within label 
  - within label variable importance may be different 
- build model for each label/drug 
- random forest LOO CV dom
- iterate by narrowing in on high importance genes? 


```{r}

temp <- merge(exp.dat, train.dat, by='cell.line')
dim(temp)

t2 <- temp[!(temp$cell.line %in% train.dat$cell.line)]
```



