---
title: 'Model # 2'
author: "Nathaniel Evans"
date: "March 15, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # all the good stuff
library(readxl) # for reading in xlsx files
library(janitor) # for clean_names
library(knitr) # for kable
library(moderndive) # for getting tables
library(corrr) # for correlation matrix
library(skimr) # for skim
library(GGally) # for ggpairs
library(broom) # for pulling out model results
library(dplyr)
library(ggpubr)
library(randomForest)

select <- dplyr::select


library(tidyr)
library(dplyr) 
library(ggplot2) 
library(moments)
library(Rtsne)

library(e1071)
```



```{r}

load('./data/train.rdat')
load('./data/exp.rdat')
load('./data/idmap.rdat')

cell.line <- as.character(exp.dat$cell.line)
exp.dat <- exp.dat[-1]
gene.names <- colnames(exp.dat)
feat.names <- character(length= length(gene.names))
for (i in 1:length(gene.names)) { 
    feat.names[i] <- paste('x', i, sep='')
}

colnames(exp.dat) <- feat.names
feat.map <- cbind(gene.names, feat.names)

train.dat <- train.dat %>% mutate(cell.line = as.character(cell.line))
train.dat$cell.line[1] <- 'X184A1'

head(feat.map)
head(exp.dat[,1:5])
head(train.dat)

```

# variance filter

Plot variance and look for elbow. Purpose is to remove genes that are displaying normal variance, ie, no expected seperation due to subtype differences. 

```{r}

skew <- numeric(length=length(colnames(exp.dat)))
kurt <- numeric(length=length(colnames(exp.dat)))
variance <- numeric(length=length(colnames(exp.dat)))
for (i in 2:length(colnames(exp.dat))) {
  variance[i] <- var(exp.dat[,i])
  skew[i] <- skewness(exp.dat[,i])
  kurt[i] <- kurtosis(exp.dat[,i])
}

s1 <- data.frame(variance, skew, kurt)

s1 %>% summary()
s1 <- s1 %>% filter (variance <3)
s1 %>% ggplot(aes(x=variance)) + geom_density() + geom_vline(xintercept = 0.35, color='red')

hivar.exp <- exp.dat[(s1$variance > 0.35)]

```


# PCA down to 10 variables 

```{r}
test <- hivar.exp[1:5, 1:5]

t(test)
```


```{r}
trans.hivar.exp <- t(hivar.exp)

colnames(trans.hivar.exp) <- cell.line

pca <- prcomp( ~ . , data=data.frame(trans.hivar.exp), center=T, scale=T, rank=10)

plot(pca)

pca.exp <- pca$rotation
```

# manually observe pc variation

```{r}
#head(train.dat)

drug <- train.dat %>% select(cell.line, CGC.11047)
pca.exp2 <- data.frame(pca.exp) %>% mutate(cell.line = cell.line)
pca.exp3 <- pca.exp2 %>% merge(drug, by='cell.line') %>% mutate(CGC.11047 = as.factor(CGC.11047))

pca.exp3 %>% head()

pca.exp3 %>% ggplot(aes(x=PC1, y=PC2, color=CGC.11047)) + geom_point()
pca.exp3 %>% ggplot(aes(x=PC2, y=PC3, color=CGC.11047)) + geom_point()
pca.exp3 %>% ggplot(aes(x=PC3, y=PC4, color=CGC.11047)) + geom_point()
pca.exp3 %>% ggplot(aes(x=PC4, y=PC5, color=CGC.11047)) + geom_point()
pca.exp3 %>% ggplot(aes(x=PC5, y=PC6, color=CGC.11047)) + geom_point()
pca.exp3 %>% ggplot(aes(x=PC6, y=PC7, color=CGC.11047)) + geom_point()
pca.exp3 %>% ggplot(aes(x=PC7, y=PC8, color=CGC.11047)) + geom_point()
pca.exp3 %>% ggplot(aes(x=PC8, y=PC9, color=CGC.11047)) + geom_point()
pca.exp3 %>% ggplot(aes(x=PC9, y=PC10, color=CGC.11047)) + geom_point()


```

```{r}
kern = 'radial'

dat <- pca.exp3 %>% mutate( y = CGC.11047 )  %>% select(-CGC.11047, -cell.line)

tc <- tune.control(random = T, nrepeat = 10, repeat.aggregate = mean,
sampling = 'cross', sampling.aggregate = mean,
sampling.dispersion = sd,
cross = 2, best.model = F,
performances = TRUE, error.fun = NULL)


params <- tune(method = svm, train.x = y ~ ., data = dat, ranges =
list(cost = seq(0.001,10,0.5), gamma=2^(-5:1)), kernel = 'radial', tunecontrol = tc)

plot(params)

params
```
svm.model <- svm(y ~., data = dat, kernel  =
kern, cost = params$cost, degree = params$degree, gamma = params$gamma, coef.0 = params$coef0,
epsilon = params$epsilon)

plot(svm.model, dat)


```{r}

# inputs: 
  # feats - 10 PC features (+cell.line) and 39 obs
  # train - cell.line, drug and 25 obs
# outputs
  # predictions - cell.line, prediction [0,1] and 16 obs 

svm.model.predictor <- function(feats, train) { 
  drug.name <- colnames(train)[2]
  colnames(train) <- c('cell.line', 'y')
  train <- train %>% mutate(y = as.factor(y))
  tr.dat <- merge(train,feats, by='cell.line')
  tr.lines <- tr.dat$cell.line
  tr.dat <- tr.dat %>% select(-cell.line)
  
  tc <- tune.control(random = T, nrepeat = 10, repeat.aggregate = mean,
  sampling = 'cross', sampling.aggregate = mean,
  sampling.dispersion = sd,
  cross = 2, best.model = F,
  performances = TRUE, error.fun = NULL)
  
  params <- tune(method = svm, train.x = y ~ ., data = tr.dat, ranges =
  list(cost = seq(0.001,10,1), gamma=2^(-5:1)), kernel = 'radial', tunecontrol = tc)
  
  cost <- params$best.parameters$cost
  gamma <- params$best.parameters$gamma
  
  svm.model <- svm(y ~., data = tr.dat, kernel  = 'radial', cost = cost, gamma = gamma)
  train.preds <- data.frame(predictions = as.numeric(as.character(predict(svm.model, (tr.dat %>% select(-y)) ))))
  train.acc <- sum(as.numeric(as.character(tr.dat$y)) == train.preds) / 25
  print(paste(drug.name, ' training accuracy: ', train.acc, '  2-fold cv acc (via tune): ', (1-params$best.performance)))
  
  test.dat <- feats %>% filter( !(cell.line %in% tr.lines) ) 
  tst.lines <- test.dat$cell.line
  test.dat <- test.dat %>% select(-cell.line)
  
  preds <- data.frame(predictions = as.numeric(as.character(predict(svm.model, test.dat))))
  
  line.preds <- cbind(data.frame(cell.line = tst.lines), preds)
  
  return(line.preds)
}


```

```{r}

#drug <- train.dat %>% select(cell.line, CGC.11047)
#pca.exp2 <- data.frame(pca.exp) %>% mutate(cell.line = cell.line$cell.line)
#pca.exp3 <- pca.exp2 %>% merge(drug, by='cell.line') %>% mutate(CGC.11047 = as.factor(CGC.11047))

train.dat2 <- train.dat %>% select(-subtype)

for (i in 1:12) { 
  drg <- colnames(train.dat2)[(i+1)]
  train1 <- train.dat %>% select(cell.line, drg) 
  
  prds <- svm.model.predictor(feats=pca.exp2, train = train1) %>% mutate(drug = drg)
  
  if (i == 1) { 
    all.prds <- prds  
  } else { 
    all.prds <- rbind(all.prds, prds)  
  }

}
```


```{r}

temp <- as.character(id.map$cell.line)
for (i in 1:168) { 
  if (temp[i] %in% '600MPE') { 
    temp[i] <- 'X600MPE'  
  }
}
id.map$cell.line <- temp

temp <- as.character(id.map$drug)
for (i in 1:168) { 
  if (temp[i] %in% 'CGC-11047') { 
    temp[i] <- 'CGC.11047'  
  } 
  if (temp[i] %in% "PF-3084014") {
    temp[i] <- 'PF.3084014'
  } 
  if (temp[i] %in% "PF-3814735") {
    temp[i] <- "PF.3814735"
  } 
  if (temp[i] %in% "PF-4691502") {
    temp[i] <- "PF.4691502"
  } 
}
id.map$drug <- temp

all.prds$cell.line <- as.character(all.prds$cell.line)

# setdiff(x,)

sub.dat <- merge(id.map, all.prds, by=c('drug', 'cell.line')) %>% mutate(value = predictions) %>% select(id, value) %>% arrange(id)

write.table(sub.dat, file='./evans_submission.csv', sep=',', row.names = F)
```


```{r}
all.prds$drug

id.map$drug
```






